I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-10 Paper/Paper_noProof.tex
fixed bug

Consider the simpler abstraction, called the {\em box abstraction} \cite{deeppoly}: from the bounds in the input, one compute by the linear step the bound on the next layer, apply the ReLU function, and so on. For instance, on the DNN example of Fig \ref{fig1}, $x_1,x_2 \in [-1,1]$, then $x_3,x_4 \in [-2,2]$, using ReLU $X_5,x_6 \in [0,2]$, and $x_7 \in [0,6]$ while $x_8 \in [1.5,3.5]$. While the bounds of $x_1, \ldots, x_6$ are correct (for all $\alpha$ 
in the range, there is an input $\vy$ such that $weight_{\vy}(x_i))=\alpha$), 
it is no more the case from the next layer (starting in $x_7,x_8$) because of potential 
dependencies between earlier neurons: $x_3$ reaches weight $2$ only when $x_1=x_2=2$,
in which case $x_4$ reaches weight $0$. Hence, it is impossible that $weight(x_7)$ reaches $6$, as it would require both $x_3$ and $x_4$ to have weight $2$.

A still extremely fast algorithm that aleviates some of this issue is known as {\em DeepPoly} \cite{DeepPoly}, also rediscovered as the CROWN algorithm \cite{Crown}. Instead of 
hard bounds, for each neuron $n$ of layer $k$, two affine functions (with input at the previous layer $k-1$) are kept as lower bound and upper bound of the weight of the node.
For instance, $f_i \leq weight(x_i) \leq g_i$ with $f_5(x_3)=f_6(x_4)=0$ and 
$g_5(x_3) = \frac{x_3+2}{2}$,
$g_6(x_4) = \frac{x_4+2}{2}$. 
With that, we get 
$x_7 \leq g_5(x_3) + 2 g_6(x_4) = \frac{x_3 + 2x_4 + 6}{2} = \frac{2x_1 - x_4 + 6}{2}\leq 4.5$.

@@

\subsection{MILP formation of $\ReLU$ functions}

In literature, many methods have been used to approximate $\ReLU$ functions. One of them is MILP (or MIP, Mixed Integer Linear Programming).

For a activation $x=\ReLU(y)$, if we already know $y's$ rough upper and lower bounds $u,l$, then if $u>0$ and $l<0$, we call $y$ is unstable; otherwise, $y$ is stable. In practice, stable node is easier because $\ReLU$ function becomes a linear function on this node. The key problem is to deal with unstable nodes.

>>>>

First, remark that weighted sums is a linear function, so it is easy to represent it explicitly. The only problematic part to represent accurately is the ReLU function. While it is fairly simple piecewise linear function with 2 modes ($x<0$ and $x \geq 0$), it is not linear. Considering its iterated impact along the different layers of a ReLU DNN is much more complex that it seems. First, let us remark that it is easy to represent exactly $ReLU(x)$
in case $x$ is known to be entirely positive and in the case where it is entirely negative
(such a neuron $x$ is called stable), as in both case there is only one known linear mode involved. Hence, the sole problem remains how to treat unstable neurons.

Consider the simpler abstraction, called the {\em box abstraction} \cite{deeppoly}: from the bounds in the input, one compute by the linear step the bound on the next layer, apply the ReLU function, and so on. For instance, on the DNN example of Fig \ref{fig1}, $x_1,x_2 \in [-1,1]$, then $x_3,x_4 \in [-2,2]$, using ReLU $X_5,x_6 \in [0,2]$, and $x_7 \in [0,6]$ while $x_8 \in [1.5,3.5]$. While the bounds of $x_1, \ldots, x_6$ are correct (for all $\alpha$ 
in the range, there is an input $\vy$ such that $weight_{\vy}(x_i))=\alpha$, 
it is no more the case from the next layer (starting in $x_7,x_8$) because of potential 
dependencies between earlier neurons: $x_3$ reaches weight $2$ only when $x_1=x_2=2$,
in which case $x_4$ reaches weight $0$. Hence, it is impossible that $weight(x_7)$ reaches $6$, as it would require both $x_3$ and $x_4$ to have weight $2$.

A still extremely fast algorithm that aleviates some of this issue is known as {\em deeppoly} \cite{deeppoly}, also rediscovered as the CROWN algorithm \cite{crown}. Instead of 
hard bounds, for each neuron $n$ of layer $k$, two affine functions (with input at the previous layer $k-1$) are kept as lower bound and upper bound of the weight of the node.
For instance, $f_i \leq weight(x_i) \leq g_i$ with $f_5(x_3)=f_6(x_4)=0$ and 
$g_5(x_3) = \frac{x_3+2}{2}$,
$g_6(x_4) = \frac{x_4+2}{2}$. 
With that, we get 
$x_7 \leq g_5(x_3) + 2 g_6(x_4) = \frac{x_3 + 2x_4 + 6}{2} = \frac{2x_1 - x_4 + 6}{2}\leq 4.5$.

@@

\subsection{MILP formulation of DNNs}

At the other end of the spectrum, we find the Mixed Integer Linear Programming (MILP) value abstraction, which is a priori a complete method:

### 2024-01-10 Paper/introduction.tex
fixed bug

For comparison, $\alpha$-$\beta$ CROWN \cite{} (and other Branch and Bound algorithms, such as \cite{BaB}) will run few instances of branch and bound (one per output neuron), in worst case considering all the possible ReLU configurations (although the branch and bound algorithm avoids most of the possibilities). On simple networks, such as those trained robust, 
this is particularly efficient because branch and bound can find very efficiently the bounds
focusing on the actual question, considering the important branchs, while our algorithm will be less efficient as it has to consider each nodes one by one from the start.
However, branch and bound faces a complexity wall when the network is hard to verify, such as the DNNs trained naturally, as there are too many branches to consider and not enough can be discarded early.

>>>>

For comparison, $\alpha$-$\beta$ CROWN \cite{crown} (and other Branch and Bound algorithms, such as \cite{BaB}) will run few instances of branch and bound (one per output neuron), in worst case considering all the possible ReLU configurations (although the branch and bound algorithm avoids most of the possibilities). On simple networks, such as those trained robust, 
this is particularly efficient because branch and bound can find very efficiently the bounds
focusing on the actual question, considering the important branchs, while our algorithm will be less efficient as it has to consider each nodes one by one from the start.
However, branch and bound faces a complexity wall when the network is hard to verify, such as the DNNs trained naturally, as there are too many branches to consider and not enough can be discarded early.
